---
title: "Case Study 10"
author: Willow Flood
date: November 8, 2022
output: github_document
---
#Data pre-processing
##The libraries for this case study
```{r message=FALSE}
library(raster)
library(rasterVis)
library(rgdal)
library(ggmap)
library(tidyverse)
library(knitr)
library(ncdf4)
```
##Creating a folder to hold the downloaded data
```{r}
dir.create("data",showWarnings = F) #create a folder to hold the data
```
##Downloading the data
```{r message=FALSE}
lulc_url="https://github.com/adammwilson/DataScienceData/blob/master/inst/extdata/appeears/MCD12Q1.051_aid0001.nc?raw=true"
lst_url="https://github.com/adammwilson/DataScienceData/blob/master/inst/extdata/appeears/MOD11A2.006_aid0001.nc?raw=true"

# download them
download.file(lulc_url,destfile="data/MCD12Q1.051_aid0001.nc", mode="wb")
download.file(lst_url,destfile="data/MOD11A2.006_aid0001.nc", mode="wb")
```
Loading the data into R:
```{r message=FALSE, warning = FALSE}
#Any warnings can be ignored
lulc=stack("data/MCD12Q1.051_aid0001.nc",varname="Land_Cover_Type_1")
lst=stack("data/MOD11A2.006_aid0001.nc",varname="LST_Day_1km")
```
##Exploring the Land Use Land Cover data
```{r}
plot(lulc)
```
Picking one year to work with to make things simpler to work with:
```{r}
lulc_13 = lulc[[13]]
plot(lulc_13)
```
##Processing the Landcover data
The land cover classes are assigned from the MODIS website
```{r}
Land_Cover_Type_1 = c(
    Water = 0, 
    `Evergreen Needleleaf forest` = 1, 
    `Evergreen Broadleaf forest` = 2,
    `Deciduous Needleleaf forest` = 3, 
    `Deciduous Broadleaf forest` = 4,
    `Mixed forest` = 5, 
    `Closed shrublands` = 6,
    `Open shrublands` = 7,
    `Woody savannas` = 8, 
    Savannas = 9,
    Grasslands = 10,
    `Permanent wetlands` = 11, 
    Croplands = 12,
    `Urban & built-up` = 13,
    `Cropland/Natural vegetation mosaic` = 14, 
    `Snow & ice` = 15,
    `Barren/Sparsely vegetated` = 16, 
    Unclassified = 254,
    NoDataFill = 255)

lcd=data.frame(
  ID=Land_Cover_Type_1,
  landcover=names(Land_Cover_Type_1),
  col=c("#000080","#008000","#00FF00", "#99CC00","#99FF99", "#339966", "#993366", "#FFCC99", "#CCFFCC", "#FFCC00", "#FF9900", "#006699", "#FFFF00", "#FF0000", "#999966", "#FFFFFF", "#808080", "#000000", "#000000"),
  stringsAsFactors = F)
# colors from https://lpdaac.usgs.gov/about/news_archive/modisterra_land_cover_types_yearly_l3_global_005deg_cmg_mod12c1
kable(head(lcd))
```
Convert the lulc_13 to a factor (catagorical) raster.
```{r}
# convert to factor raster (easy). This creates a raster attribute table (RAT)
lulc_13=as.factor(lulc_13)

# update the RAT with a left join
levels(lulc_13)=left_join(levels(lulc_13)[[1]],lcd)
```
Plotting to see how it looks:
```{r}
gplot(lulc_13)+
  geom_raster(aes(fill=as.factor(value)))+
  scale_fill_manual(values=levels(lulc_13)[[1]]$col,
                    labels=levels(lulc_13)[[1]]$landcover,
                    name="Landcover Type")+
  coord_equal()+
  theme(legend.position = "bottom")+
  guides(fill=guide_legend(ncol=1,byrow=TRUE))
```
##Exploring the Land Surface Temperature data
```{r}
plot(lst[[1:12]])
```
Convert lst from Kelvin to Celsius:
```{r}
offs(lst)=-273.15
plot(lst[[1:10]])
```
Adding dates to Z time
```{r}
names(lst)[1:5] #default dates
```
Need to convert them to an R date and drop the "X":
```{r}
tdates=names(lst)%>%
  sub(pattern="X",replacement="")%>%
  as.Date("%Y.%m.%d")

names(lst)=1:nlayers(lst)
lst=setZ(lst,tdates)
```
#Part 1:Extract timeseries for a point
Extract LST values for a single point and plot them.
```{r}
#Spatial point to extract data from
lw = SpatialPoints(data.frame(x= -78.791547,y=43.007211)) 
#Setting the projection of the point
projection(lw) <- "+proj=longlat"
#Transforming the point to the raster projection
lw = spTransform(lw, CRSobj = crs(lst, asText = TRUE))
```
Extracting the lst data from that point:
```{r}
e = raster::extract(lst,lw,buffer=1000,fun=mean,na.rm=T)
et = t(e)
```
Extracting the dates and creating a data frame
```{r message=FALSE}
dates = getZ(lst)
lst_dates = bind_cols(et, dates)
```
Plotting the data:
```{r warning=FALSE}
ggplot(lst_dates, aes(x = ...2, y = ...1)) +
  geom_point() +
  geom_smooth(span = 0.02, n = 400) +
  labs(title = "Monthly Mean Land Surface Temperature",
       x = "Date",
       y = "Mean Land Surface Temperature")
```
#Part 2: Summarize weekly data to monthly climatologies
Creating the monthly mean variables:
```{r}
#Sub-setting the months
tmonth = as.numeric(format(getZ(lst),"%m"))
#Summarizing the mean value per month
lst_month = stackApply(lst, indices = tmonth, fun = mean)
#Setting the names of the layers to months
names(lst_month)=month.name
```
Plotting the map for each month:
```{r}
gplot(lst_month)+
  geom_raster(aes(fill = value))+
  facet_wrap(~ variable)+
  scale_fill_gradient(low='blue',high='red')+
  labs(x = "X", y = "Y")+
  theme(axis.text.x=element_blank(),axis.text.y=element_blank())
```
Calculating the monthly mean for the entire image.
```{r}
cellStats(lst_month, mean) %>% 
  as.data.frame() %>% 
  rename(Mean = ".") %>% 
  kable()
```
#Part 3: Summarize Land Surface Temperature by Land Cover
Resampling the land cover data.
```{r}
lulc2 <- resample(lulc, lst, method="ngb")
```
Extracting the values from lst_month and lulc2, gathering the data, converting the ID to numeric, joining the data to the data frame, filtering the data to the specifications.
```{r}
lcds1=cbind.data.frame(
  values(lst_month),
  ID=values(lulc2[[1]]))%>%
  na.omit() %>%
  gather(key='month', value='value', -ID) %>%
  mutate(ID=as.numeric(ID)) %>%
  mutate(month=factor(month, levels=month.name, ordered=T)) %>%
  left_join(lcd) %>%
  filter(landcover%in%c("Urban & built-up","Deciduous Broadleaf forest"))
```
Plotting the new data.
```{r}
ggplot(lcds1)+
  geom_point(aes(month,value))+
  theme_bw()+
  facet_wrap(~landcover)+
  labs(y = "Monthly Mean Land Surface Temperature (C)")
```
